Enabling the Use of GPU Virtualization in Cloud Environments

### 2015-10-08 paper.tex
ispell pasado

Thanks to the shareability introduced by {rCUDA}, all VM in the cloud 
environment should be able to use and share all the GPUs installed in 
the physical nodes removing the current limitation. But, in order to achieve this goal, 
the cloud facilities need to be provided with the {rCUDA} software and 
in the case of OpenStack, a module for manageing virtualized GPUs 
 has to be implemented.

@@

For this study, carried out in June 2015, we worked in the AWS availability zone US EAST (N. VIRGINIA). 
So, at this moment the instances used had the features presented in the Table~\ref{table:awsInstances}.

@@

The selection of these concrete types have the aim of using C3 family instances, which are not equipped with GPUs, as client whereas the instances of the G2 family are acting as servers.

@@

The IPERF tool has been executed between the instances showed in 
the Table~\ref{table:iperf} in order to know the real bandwidth.
Therefore, we can assure that the nomenclature ``High'' used in the 
network column correspond to a 1~Gb interconnection and the 10~Gb has 
a real bandwidth of 7.5~Gb/s.
Moreover, it looks like the instances equipped with a ``High'' interconnection 
are software-limited up to 1~Gb/s due to the theoretical and real bandwith 
match perfectly. The gap between real and theorical can be observed with 
the 10~Gb interconnection which reaches up to 7.5~Gb/s.

\subsubsection{GPUs}
As it has been said before, an instance is based in a VM which is running on a real node with its own virtualized components, so it is possible that AWS could use a virtualization framework to offer GPU services to all the instances.
Although, the {\tt nvidia-smi} command informs us that the GPUs installed are NVIDIA GRID K520, we need to verify that they are non-virtualized devices so, we have executed the NVIDIA SDK test called {\tt bandwidthtest}. 
As it is shown in the Table~\ref{table:bwt}, as the bandwidth achieved in this test is higher than the network bandwith, it reveals that it is a real GPU. 
\begin{table}[htb]
\renewcommand{\arraystretch}{1.3}
\caption{NVIDIA SDK {\tt bandwidthtest} execution transfering 32MB using pageable memory using local GPU.}
\label{table:bwt}
\tabcolsep=0.09cm
\begin{center}\begin{tabular}{cccc}
Name &  Data Movement & Network & BandWidth \\ \hline \hline
g2.2xlarge & Host to Device & High& 3,004 MB/s \\ \hline
g2.2xlarge & Device to Host & High& 2,809 MB/s\\ \hline
g2.8xlarge & Host to Device & 10 Gb& 2,814 MB/s\\ \hline
g2.8xlarge & Device to Host & 10 Gb& 3,182 MB/s\\ \hline
\end{tabular}\end{center}\end{table}

@@

\item Scenario B (Figure~\ref{subfig:aws2}) is composed by 2 server nodes equipped 
with 4 GPUs each one and a client that does not have any GPU. This scenario is using 
a 10Gb network and the cliend is able to execute application using up to 8 GPUs.

@@

\begin{table}[htb]
\renewcommand{\arraystretch}{1.3}
\caption{NVIDIA SDK {\tt bandwidthtest} execution transfering 32MB using pageable memory using {rCUDA}.}
\label{table:bwtrcuda}
\tabcolsep=0.09cm
\begin{center}\begin{tabular}{cccc}
Scenario &  Data Movement & Network & BandWidth \\ \hline \hline
A & Host to Device & High& 127 MB/s \\ \hline
A & Device to Host & High& 126 MB/s\\ \hline
B & Host to Device & 10 Gb& 858 MB/s\\ \hline
B & Device to Host & 10 Gb& 843 MB/s\\ \hline
\end{tabular}\end{center}\end{table}

@@

\subsection{Discussion}
Putting all together,  it is cristal clear that the AWS GPU-instances are not aimed to HPC because neither the network nor the accelerators are powerful enough to reach an appropriate performance
while running intensive parallel applications. 
Hence, AWS's target is more oriented to offer the service than to provide performance.
Also with the aim of plainly offering the service of GPGPU, AWS lacks in flexibility. 
AWS forces the user to choose between a basic instance with a GPU and a powerful instance with 4 GPUs.

@@

\subsection{Managing remote GPUs with OpenStack}
The main idea is to move from the original Openstack architecture (see Subfigure~\ref{subfig:os-orig}) to a solution where a new shared service takes part into the architecture and is responsible for all related to GPU accelerators (see Subfigure~\ref{subfig:os-mod}).
This new service would bring more flexibility when managing GPUs and new working modes for GPGPU computation in the Cloud.
As it is shown in Figure~\ref{fig:internal}, we have altered the OpenStack original Dashboard with a new parser, 
which split the HTTP query in order to make use of both, the GPGPU API for everything related to GPUs, and the nova API for the rest of the computation. 
The new GPGPU Service has brought a new feature which brings more flexibility when managing GPUs, from now on, we can create ``GPU-pools''. 
A ``GPU-pool'' is a set of independent GPUs (logically disattached from the nodes) that can be assigned to one or more instances.

\begin{figure}[htb]
  \centering
  \subfigure[Original in version XXXXXXXXXXXX]{
    \includegraphics[width=\linewidth]{images/os-orig.png}
    \label{subfig:os-orig}
   }
   \quad
  \subfigure[With GPGPU developed module]{
    \includegraphics[width=\linewidth]{images/os1.jpg}
    \label{subfig:os-mod}
   }
  \caption{Openstack Architecture}
  \label{fig:os}
\end{figure}

@@

Finally, to bring in more strategies to decide where a remote GPUs is created and assigned to a physical device, would be an interesting task to carry out. 
With more scheduling policies the flexibility given by the GPGPU module for Openstack would increase.

>>>>

Thanks to the shareability introduced by {rCUDA}, all VM in the cloud 
environment should be able to use and share all the GPUs installed in 
the physical nodes removing the current limitation. But, in order to achieve this goal, 
the cloud facilities need to be provided with the {rCUDA} software and 
in the case of OpenStack, a module for managing virtualized GPUs 
 has to be implemented.

@@

For this study, the instances presented in the Table~\ref{table:awsInstances} 
have been selected. The C3 family instances, which are not equipped with 
GPUs are going to be used as client whereas the instances of the G2 family are 
acting as servers.

@@

@@

The IPERF tool has been executed between the instances showed in 
the Table~\ref{table:iperf} in order to know the real bandwidth.
Therefore, we can assure that the nomenclature ``High'' used in the 
network column correspond to a 1~Gb interconnection and the 10~Gb has 
a real bandwidth of 7.5~Gb/s.
Moreover, it looks like the instances equipped with a ``High'' interconnection 
are software-limited up to 1~Gb/s due to the theoretical and real bandwidth 
match perfectly. The gap between real and theoretical can be observed with 
the 10~Gb interconnection which reaches up to 7.5~Gb/s.

\subsubsection{GPUs}
As it has been said before, an instance is based in a VM 
which is running on a real node with its own virtualized 
components, so it is possible that AWS could use a virtualization 
framework to offer GPU services to all the instances.
Therefore, we need to verify if the GPUs showed in the instances are 
non-virtualized GPUs so, we have executed the NVIDIA SDK test called 
{\tt bandwidthtest}. As it is shown in the Table~\ref{table:bwt}, as the bandwidth achieved 
in this test is higher than the network bandwidth, it reveals that it is a 
real GPU. 
\begin{table}[htb]
\renewcommand{\arraystretch}{1.3}
\caption{NVIDIA SDK {\tt bandwidthtest} execution transferring 32MB using pageable memory using local GPU.}
\label{table:bwt}
\tabcolsep=0.09cm
\begin{center}\begin{tabular}{cccc}
Name &  Data Movement & Network & Bandwidth \\ \hline \hline
g2.2xlarge & Host to Device & High& 3,004 MB/s \\ \hline
g2.2xlarge & Device to Host & High& 2,809 MB/s\\ \hline
g2.8xlarge & Host to Device & 10 Gb& 2,814 MB/s\\ \hline
g2.8xlarge & Device to Host & 10 Gb& 3,182 MB/s\\ \hline
\end{tabular}\end{center}\end{table}

@@

\item Scenario B (Figure~\ref{subfig:aws2}) is composed by 2 server nodes equipped 
with 4 GPUs each one and a client that does not have any GPU. This scenario is using 
a 10Gb network and the client is able to execute application using up to 8 GPUs.

@@

\begin{table}[htb]
\renewcommand{\arraystretch}{1.3}
\caption{NVIDIA SDK {\tt bandwidthtest} execution transferring 32MB using pageable memory using {rCUDA}.}
\label{table:bwtrcuda}
\tabcolsep=0.09cm
\begin{center}\begin{tabular}{cccc}
Scenario &  Data Movement & Network & Bandwidth \\ \hline \hline
A & Host to Device & High& 127 MB/s \\ \hline
A & Device to Host & High& 126 MB/s\\ \hline
B & Host to Device & 10 Gb& 858 MB/s\\ \hline
B & Device to Host & 10 Gb& 843 MB/s\\ \hline
\end{tabular}\end{center}\end{table}

@@

\subsection{Discussion}
Putting all together,  it is crystal clear that the AWS GPU-instances are not aimed to HPC because neither the network nor the accelerators are powerful enough to reach an appropriate performance
while running intensive parallel applications. 
Hence, AWS's target is more oriented to offer the service than to provide performance.
Also with the aim of plainly offering the service of GPGPU, AWS lacks in flexibility. 
AWS forces the user to choose between a basic instance with a GPU and a powerful instance with 4 GPUs.

@@

\subsection{Managing remote GPUs with OpenStack}
The main idea is to move from the original OpenStack architecture (see Subfigure~\ref{subfig:os-orig}) to a solution where a new shared service takes part into the architecture and is responsible for all related to GPU accelerators (see Subfigure~\ref{subfig:os-mod}).
This new service would bring more flexibility when managing GPUs and new working modes for GPGPU computation in the Cloud.
As it is shown in Figure~\ref{fig:internal}, we have altered the OpenStack original Dashboard with a new parser, 
which split the HTTP query in order to make use of both, the GPGPU API for everything related to GPUs, and the nova API for the rest of the computation. 
The new GPGPU Service has brought a new feature which brings more flexibility when managing GPUs, from now on, we can create ``GPU-pools''. 
A ``GPU-pool'' is a set of independent GPUs (logically disattached from the nodes) that can be assigned to one or more instances.

\begin{figure}[htb]
  \centering
  \subfigure[Original in version XXXXXXXXXXXX]{
    \includegraphics[width=\linewidth]{images/os-orig.png}
    \label{subfig:os-orig}
   }
   \quad
  \subfigure[With GPGPU developed module]{
    \includegraphics[width=\linewidth]{images/os1.jpg}
    \label{subfig:os-mod}
   }
  \caption{OpenStack Architecture}
  \label{fig:os}
\end{figure}

@@

Finally, to bring in more strategies to decide where a remote GPUs is created and assigned to a physical device, would be an interesting task to carry out. 
With more scheduling policies the flexibility given by the GPGPU module for OpenStack would increase.
